Halving the cost of quantum addition

### 2018-05-25 paper.tex

It is possible to produce a high-quality $|T\rangle$ state in 960 units of ``spacetime volume" \cite{} (though this is certainly not a lower bound!).
For comparison, an ancilla qubit stored as a double-defect through $k$ serial measurements will cover at least $2k$ units of spacetime volume.
Putting these two facts together, and ignoring the fact that T factories are discrete objects with packing constraints, we will approximate the opportunity cost of holding an ancilla as $\frac{1}{480} |T\rangle$ states per measurement-depth.

@@

We caution that the cutoff of 960 that we just estimated is not a fixed constant.
It is affected by other optimizations, by future research, and by overall system design.
As the spacetime volume of T factories is improved, the cutoff {\em will} move downward.
Conversely, if the spacetime volume of ancillae is improved, the cutoff moves upward.
For example, qubits that are idle can be compressed into an inactive ``memory" form that covers six times fewer physical qubits than the double-defect form \cite{horsman2012}.
This means that {\em idle} ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
This optimization increases the cutoff to $n \approx 5760$, which is beyond the size of adders that would be used to break modern RSA keys with Shor's algorithm.

@@

Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
For example, we used temporary logical-ANDs to cut the projected T-count of a step in a chemistry algorithm under development by nearly 30\% \cite{RyanEmails2017}.

>>>>

It is possible to produce a high-quality $|T\rangle$ state in 960 units of ``spacetime volume" \cite{Babbush2018} (though this is certainly not a lower bound!).
For comparison, an ancilla qubit stored as a double-defect through $k$ serial measurements will cover at least $2k$ units of spacetime volume.
Putting these two facts together, and ignoring the fact that T factories are discrete objects with packing constraints, we will approximate the opportunity cost of holding an ancilla as $\frac{1}{480} |T\rangle$ states per measurement-depth.

@@

We caution that the cutoff of 960 that we just estimated is not a fixed constant.
It is affected by other optimizations, by future research, and by overall system design.
As the spacetime volume of T factories is improved, the cutoff {\em will} move downward.
Conversely, if the spacetime volume of ancillae is improved, the cutoff moves upward.
For example, qubits that are idle can be compressed into an inactive ``memory" form that covers six times fewer physical qubits than the double-defect form \cite{horsman2012}.
This means that {\em idle} ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
This optimization increases the cutoff to $n \approx 5760$, which is beyond the size of adders that would be used to break 4096-bit RSA keys with Shor's algorithm.

@@

Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
For example, we used temporary logical-ANDs to cut the T-count of a chemistry algorithm in preparation by nearly 50\% \cite{Babbush2018}.
